Search-Based Test Data Generation for JavaScript Functions that Interact with the DOM

### 2018-05-07 doc/icse2018/issta2018.tex
Working on issre 2018 revision.

\documentclass[sigconf,review,anonymous]{acmart}

@@

\listfiles % used to output the versions of all files being used by the latex

@@

% Copyright
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

@@

\author{Jurriaan Hage}
\affiliation{%
  \institution{Utrecht University, The Netherlands}
  \country{The Netherlands} 
}
\email{j.hage@uu.nl}

\begin{abstract}
As of 2017 JavaScript and Node.js continue to stay the most prevalent technologies according to the Stack Overflow developers survey. Nevertheless, existing JS testing frameworks are still lacking any support for such a vital step in test automation as the generation of input data. To address this problem, we present a novel \textbf{J}avascript \textbf{E}volutionary-based testing framework with \textbf{D}OM as an \textbf{I}input, called \textbf{JEDI}. In order to reach a target branch, it applies genetic search for relevant input parameters of the JS function including global DOM state. We conducted an empirical evaluation to study the effectiveness and efficiency of our testing framework.
 It has shown that the \emph{genetic with restart} algorithm, proposed in this paper, was able to achieve complete branch coverage on all experimental subjects with the average 19 seconds of execution time per branch.
\end{abstract}

@@

\section{Introduction}
\label{sec.intro}

According to a Stack Overflow survey\footnote{\url{https://insights.stackoverflow.com/survey/2017}} run in 2017 across developers, JavaScript has been recognized for the fifth time in row as the most commonly used programming language with Node.js as the most commonly used technology. It is used through all technology stack web, back-ends and mobile. Such a tremendous ever increasing JS success is happening thanks to the extensive dynamic features of the language, JS permissiveness and native browser support. On the other hand, JS applications are error-prone~\cite{why}, hard to analyze~\cite{why} and test~\cite{why}. Therefore, it is not surprising that year after year we hear new stories\footnote{\url{https://www.fastcompany.com/40505226/the-year-that-software-bugs-ate-the-world}} about another critical bug being found in a web application by Facebook, Twitter etc.

%% Common mistakes JS developers do\footnote{\url{https://www.toptal.com/javascript/10-most-common-javascript-mistakes}}.

%% Notable errors in 2017\footnote{\url{https://www.fastcompany.com/40505226/the-year-that-software-bugs-ate-the-world}}.

The state of JS 2017 survey~\footnote{\url{https://stateofjs.com/}} contains an extensive list of testing frameworks including Mocha, Jasmine, Jest, Enzyme, Tape, Ava, Karma and QUnit. These frameworks support testing at various levels: unit, integration and functional. The survey also indicated relatively low satisfaction rate of the current state of JS testing tools, scoring only 3.2 out of 5. As result, new testing frameworks such as Cypress and Storybook are constantly appearing to improve the situation. But despite of this effort, test data generation continues to be not properly addressed by any of the widely used testing frameworks. The research community has actively been working towards making progress in this direction introducing such tools as Artemis~\cite{artemis2011}, JSConTest~\cite{heidegger2010contract}, Crawljax~\cite{mesbah2012crawling}, JSeft~\cite{mirshokraie2015jseft} and Jalangi~\cite{sen2013jalangi}. A recent study of client-side JS bugs~\cite{ocariza2017study} has shown that the majority (68\%) of JS faults are DOM-related and occur in JS code, and not in HTML or CSS. This fact indicates that yet a better testing approach is needed when it comes to unit testing of JS.

Over the past two decades search-based test data generation has proven to be one of the most effective solutions for a wide range of technologies including object-oriented~\cite{paper}, web~\cite{paper} and mobile~\cite{mao2016sapienz}. In recent years, some effort was made to apply search-based testing to dynamic languages such as Python~\cite{paper}, Ruby~\cite{paper} (maybe mention Go and Lua), but to the best of our knowledge search-based testing approach has never been applied to JavaScript. Our work makes the following contributions:
\begin{enumerate}[leftmargin=5mm]
\item The \emph{JEDI} framework is a novel JS unit testing tool using search-based techniques for generating test data. Notably, our testing framework is able to generate an arbitrary DOM input, which is syntactically valid and compliant with the recent HTML5 specification.   
\item We propose a novel test generation algorithm, called ``genetic with restart'', combining the benefits of systematic fitness improvement with the genetic search and the ability of escaping plateau by restarting search with a new target.
\item Based on the case studies found in two related works~\emph{Confix}~\cite{amin:ase15} and \emph{TAJS}~\cite{dom2011}, we conduced an empirical validation of our framework and performed a significance study of the results. The validation has shown the effectiveness of the framework in covering target branches with a reasonable efficiency...  
\end{enumerate}

The paper is structured as follows: In Section~\ref{sec.example}, we present a motivating example illustrating the challenges of generating test data for JS. The architecture of the \emph{JEDI} testing framework is explained in Section~\ref{sec.framework}. Section~\ref{sec.evaluation} present empirical validation, whereas Section~\ref{sec.disc} discusses future work and validity threats. We discuss related work in Section~\ref{sec:related.work} and conclude the paper in Section~\ref{sec:concl}.

Another empirical work~\cite{richards2010analysis}.

@@

Let us consider the JS function \texttt{isGameFinished} in Figure~\ref{code.isGameFinished} taken from a web game \emph{sudoku}\footnote{\url{http://www.dhtmlgoodies.com/scripts/game_sudoku/game_sudoku.html}}. This function checks the validity of a sudoku solution given by a user. First, we find a DOM element representing the game field (line~\ref{isGameFinished.getSudoku}). Then, we collect all its child DIVs (line~\ref{isGameFinished.getDivs}). Each DIV element corresponds to a sudoku input square (line~\ref{isGameFinished.if1.begin}) and consists of two SPANs. First span contains a user input, whereas the second one, which is hidden by default, stores the expected value for that sudoku square. Inside of a for-loop in lines~\ref{isGameFinished.inFor.begin}-\ref{isGameFinished.outFor}, we iterate over the whole collection of input squares. Once, we find a square with two unequal SPAN values (line~\ref{isGameFinished.if2.begin}), the loop is terminated and the game is considered unfinished (line~\ref{isGameFinished.unfinished}) and finished otherwise.

Suppose now that we would like to unit test this function in isolation to the rest of the application. We aim to maximize the branch coverage as a common testing criteria~\cite{why.this.common}. That is, for each individual branch of the function under test (FUT) \texttt{isGameFinished}, we need to construct a test input that covers the branch and leads to the normal termination of the FUT. JS function takes \emph{explicit} input arguments (in our case none), but it can also accept \emph{implicit} arguments such as the DOM state. We assume that DOM is the only global variable that can be referenced in the body of the FUT. So in order to produce a complete test for our function we have to construct both an appropriate DOM and input arguments.

@@

Figure~\ref{fig.isGameFinished.tests} presents four tests that together deliver the full coverage of the FUT. Let us look now at how following the structure of the FUT in Figure~\ref{code.isGameFinished}, we can construct the $T4$ test covering the $(10,11)$ branch. Line~\ref{isGameFinished.getSudoku} expects the input DOM to have an element with the id \textquotesingle\texttt{sudoku}\textquotesingle. In order to enter the for-loop in line~\ref{isGameFinished.inFor.begin}, that element should have at least one child \emph{DIV} tag. The first if-condition in line~\ref{isGameFinished.if1.begin} requires this $div$ to be of the class \textquotesingle\texttt{sudokuSquare}\textquotesingle\ and also have no background color. The second if-condition (line~\ref{isGameFinished.if2.begin}) expects that the $div$ consists of two \emph{SPAN} elements whose \emph{innerHtml} values are not equal. The resulted test $T4$ is shown in Figure~\ref{fig.isGameFinished.tests} and meets all the above conditions.

As we have just seen, even for such a relatively simple example, the construction procedure of a test input is far from trivial because it requires the deep understanding of the program semantics. In practice, web developers often have to deal with the gazillion of the JS frameworks and libraries freely available on the GitHub whose code is commonly untested and poorly documented. At the same time, unit testing is the most fundamental level of testing and, at the same time, it is easiest to implement. It greatly saves the developers from unexpected regression and puts a safety belt around the application.

@@

Given a JS function $f$ and a set of branches $\mathbb{B}_f$, the ultimate goal of the testing framework is to find input arguments for $f$ that cover all branches from $B_f$. Our JS testing framework, called \emph{JEDI}, is a client-server application which under the hood uses the genetic engine for test generation. The general workflow of the framework is depicted as a sequence diagram in Figure~\ref{fig.framework.architect}. The diagram consists of the four interacting components \User, \Server, \Validator, and \Client. \Server is the key component of the whole framework. It is responsible for the \emph{Initialization Phase} and the \emph{Genetic Phase}. Both phases will be covered in details in Section~\ref{sub.sec.init.phase} and \ref{sub.sec.genetic.phase} respectively.

Each newly constructed HTML document has to be processed by the \Validator\footnote{\url{https://validator.w3.org/nu/}} which checks its syntactic correctness and reports back the set of identified errors. \Server rejects the document if any errors were revealed and generates a new one until the generated document has no errors anymore. \Client is a \emph{Node.js}\footnote{\url{https://nodejs.org/en/}} application that is mainly responsible for the execution of the JS code. We use the \emph{jsdom}\footnote{\url{https://github.com/tmpvar/jsdom}} library to model the virtual DOM and simulate the native browser API calls.

@@

Let us give a closer look at the sequence of interactions in Figure~\ref{fig.framework.architect}. As an input for our testing framework, \User has to provide the JS file \emph{fun.js} that contains the function under test (FUT) \emph{fun} annotated with the type signature. Additionally, \User specifies a configuration file \emph{config} that defines various parameters for random data generation, genetic phase, logging etc. Experimentally, we have found an optimal set of input configurations that was used across all experimental objects in Section~\ref{sec.evaluation}. This data is processed by the \Server during the \emph{Initialization Phase}. As an output, we get back the following information about \emph{fun}: the control flow graph (\emph{cfg}), the complete list of its branches (\emph{branches}), the constant info (\emph{cinfo}), the instrumented version \emph{ifun}, and the signature (\emph{sig}). \Server sends a \texttt{POST} request with \emph{ifun} and \emph{sig} to the \Client. Depends on the configured execution mode, the \Server either asks the \User to select a \emph{target} branch or covers all branches one by one in a fixed order. For each branch, our framework enters the \emph{Genetic Loop (Phase)}. Inside of this phase, the framework starts with a randomly generated population (\emph{Random Loop}), and continues to evolve that population (\emph{Fitness Loop}) until either a perfect entity is found or a certain termination criteria is achieved, e.g. the limit on the number of generation attempts. Each candidate in the population has to be evaluated in \emph{Evaluation Loop} by \Client. It matches the candidate $p$ according to the type signature \emph{sig} and invokes the instrumented JS function on it: \emph{ifun(p)}. The execution trace along with some supplementary information about the execution are sent back to the \Server, where the final \emph{fitness value} is computed. The new population is constructed by combining the results of the \emph{Crossover} and \emph{Mutation Loops}. If a new HTML document is produced somewhere in the genetic phase, it has to be first approved by the \Validator. At the end of the \emph{Fitness Loop}, the candidate with the ``best'' fitness value is reported back to the \User. Once all the branches are exercised, the algorithm reaches the end. In the follow up two sections, we elaborate on the details of each individual phase of our test generation framework.

@@

During the \emph{Initialization Phase} we analyze the FUN in order to retrieve the data required for the consecutive genetic testing. The main steps of this phase are illustrated by Algorithm~\ref{alg.init}. Given a FUT \emph{fun.js} as an input, we parse the function and its type signature in line~\ref{alg.init.parse.fun.sig}. In order to maximize the chances of generating the fittest candidate in the genetic phase, in line~\ref{alg.init.collect.info} we perform static analysis of the FUT with the purpose to collect constant info. Such constants include the numeric and string literals, as well as the arguments to the following DOM API methods \texttt{getElementsByTagName}, \texttt{getElementById}, \texttt{getElementsByClassName}, and \texttt{getElementsByName},
which respectively signify the reference to the \emph{tags}, \emph{ids}, \emph{classes} and \emph{names} encountered in the FUT. Next step is normalizing the FUT before applying any instrumentation (line~\ref{alg.init.transform.ast}), e.g. transform \emph{if-then} branches into \emph{if-then-else}. We instrument the FUT (line~\ref{alg.init.instr}) in order to collect the run-time data such as the execution trace, the approach level, the branch and loop distances, and the constants. Upon the $ifun$ execution this information is gathered and fed back to the fitness evaluation procedure. Finally, we construct the control-flow graph (line~\ref{alg.init.build.cfg}) and identify all branches of the FUT (line~\ref{alg.init.get.branches}).

@@

\begin{figure}[t!]
\setlength{\grammarparsep}{3pt}
\small
\begin{grammar}

@@

The grammar of the JS type annotations supported by our test generation framework is shown in Figure~\ref{fig.js.type.annot}. Each type annotation is a signature embedded into a specialized JS type comment (\texttt{/*t...*/}). We distinguish between two kinds of signatures those which refer to the global DOM and the ``ordinary'' JS type signatures. The functions of the former signature expect some \texttt{dom} argument to be passed as the first implicit parameter for the function call. Whereas ordinary JS signature is a non-zero sequence of the JS types. The type can be \emph{Void}, \emph{Primitive} (\texttt{bool}, \texttt{int}, \texttt{float}, \texttt{string}) or \emph{Array} (recursive homogeneous array type).

@@

Figure~\ref{code.isGameFinished.instr} presents the \texttt{isGameFinished} function after instrumentation. The global variable \emph{trace} in line~\ref{isGameFinished.instr.trace} is an array of labels corresponding to the sequence of executed program statements. All statements have to be traced because each one, in theory, can rise an exception. The \texttt{loopMap} variable in line~\ref{isGameFinished.instr.loopMap} approximates the potential number of iterations for the consecutive for-loop. If no exact bound can be computed, we set it to one. The \texttt{branchDistance} map tracks the distances to the deviated branches according to the translation formulae~\cite{tracey1998automated}.

@@

The heart of our test data generation framework is in the \emph{Genetic Phase}. Though, the main workflow of the underlying genetic algorithm (GA) is fairly standard~\cite{poli2008field}, but has been extended with the concept of the \emph{archive convergence} (Algorithm~\ref{alg.gen}). As an input, we take a configuration file \emph{config}, a target branch $b_n$ of the FUT (which we would like to cover), and a tuple of supplementary parameters computed by Algorithm~\ref{alg.init}. The output consists of the ``best'' fitness score $f^*$ achieved by the population candidate $p^*$.

First, we read the \emph{config} file (line~\ref{alg.gen.read.config}) and retrieve the following GA configuration parameters: a population size  ($size_P$), an archive size ($size_A$), the number of generations ($size_G$), crossover ($rate_C$) and mutation ($rate_M$) rates respectively, and a history window size ($size_H$). We start with the initial target $b_n$ (line~\ref{alg.gen.init.target}) and the iteration counter $i$ equal to 0 (line~\ref{alg.gen.iter.count}). We run the genetic search for the target branch (lines~\ref{alg.gen.converged.loop.begin}-\ref{alg.gen.converged.loop.end}) until either the history has not converged or the generations limit has not been reached. The \emph{archive} variable stores the most fitted population candidates observed so far, and the \emph{history} variables records all archives during the generation process. These variables both are empty in the beginning of the evolution (line~\ref{alg.gen.init.archive} and~\ref{alg.gen.init.history} respectively).

>>>>

\documentclass[sigconf,review]{acmart}

@@

\listfiles % used to output the versions of all files being used by the late

@@

% Copyright
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

@@

\author{Jurriaan Hage}
\affiliation{%
  \institution{Utrecht University, The Netherlands}
}
\email{j.hage@uu.nl}

\begin{abstract}
The popularity of JavaScript is continuously growing. Together with HTML and CSS, it forms the core technology stack for web development. Because of the dynamic nature and complex interplay with HTML, JS applications are still error-prone and vulnerable. Despite the active research efforts to devise intricate static and dynamic analyses to facilitate JS testing, the problem of test data generation for JS applications interacting with the DOM has not yet been addressed. In this paper, we present a novel \textbf{J}avascript \textbf{E}volution-based testing framework with \textbf{D}OM as an \textbf{I}nput, called \emph{JEDI}. In order to reach a target branch, it applies genetic search for relevant input parameters of the JS function including the global DOM state. We conducted an empirical evaluation to study the effectiveness and efficiency of our testing framework. It shows that the \emph{genetic with restart} algorithm, proposed in this paper, is able to achieve complete branch coverage for all experimental subjects, taking on average 19 seconds per branch.
\end{abstract}

@@

%% ------------------------------------------------------------------------
\section{Introduction}
\label{sec.intro}
%% ------------------------------------------------------------------------

According to the \emph{StackOverflow} Developer Survey 2018 \cite{stackoverflow2018}, JavaScript has been recognized for the sixth time in row as the most commonly used programming language. It is also the most popular language on \emph{GitHub}~\cite{guthub2017} based on the number of opened pull requests (2.3 million). Such an ever increasing JS success can be explained by its native browsers support, and the rise in popularity of the JS run-time environment \emph{Node.js}. On the other hand, due to of implicit type conversion, lack of static types, and extensive support of dynamic language constructs~\cite{richards2010analysis}, JS applications are often error-prone~\cite{frolin:TSE16}, hard to analyze~\cite{andreasen2017survey, sun2017analysis} and test~\cite{mesbah2015advances}. Thus, it is not surprising that once in a while critical bugs manage to break through~\cite{bugstories2017} into the web applications by Facebook, Twitter etc.

Systematic unit testing is basic and, at the same time, the most effective technique for improving software quality. In practice, there are numerous of JS frameworks --- Mocha, Jasmine, Jest, Enzyme, Tape, Ava, Karma, QUnit etc. --- intended to assist developers with the automation of various unit testing activities such as design, execution and monitoring. Despite the apparent abundance of the JS testing tools, the developers' satisfaction rate is still relative low~\cite{stateJS2017} (3.2 out of 5), suggesting the need for improvements. One of such areas is \emph{test data generation}. It pursues the goal of maximizing the test coverage by exercising the function under test (FUT) with different input parameters. Moreover, if a FUT interacts with the DOM of a web application, a \emph{test fixture} is expected as an additional input.

Generally, the problem of test data generation has been under an active study for the past three decades. In particular for JS, researchers have proposed several solutions based on: concolic execution (\emph{Jalangi}~\cite{sen2013jalangi}, \emph{Confix}~\cite{amin:ase15}), random generation (\emph{JSContest}~\cite{heidegger2010contract}), and the analysis of the whole web application (\emph{Artemis}~\cite{artemis2011}, Crawljax~\cite{mesbah2012crawling}). Out of those tools, only \emph{Confix} tries to construct a sufficient DOM fixture by translating native browser API calls into logical constraints. However, in the empirical evaluation~\cite{amin:ase15}, Confix only reached about 50\% in \emph{branch coverage} on the selected subjects. At the same time, search-based techniques have shown prominent results in achieving high branch coverage for imperative~\cite{wegener2001evolutionary}, object-oriented~\cite{fraser2011evosuite} and dynamic languages~\cite{irawan2016test, wibowo2015unit}, as well as for web~\cite{alshahwan2011automated} and mobile~\cite{mao2016sapienz} applications. But to the best of our knowledge, the problem of test data generation has never been tackled for JS with the help of the search-based approach. In this paper we are breaching this gap by introducing a \textbf{J}avascript \textbf{E}volution-based testing framework with \textbf{D}OM as an \textbf{I}nput, called \emph{JEDI}. We can summarize our contributions as follows: 
\begin{enumerate}[leftmargin=5mm]
\item The \emph{JEDI} framework is a novel JS unit testing tool using search-based techniques for generating test data. Notably, our testing framework is able to generate an arbitrary DOM input (both tags and attributes), which is syntactically valid and compliant with the recent HTML5 specification.   
\item Our test generation algorithm, called ``genetic with restart'', is trying to escape search plateau by restarting GA with a new target. Because the full-fledged JS data flow analysis~\cite{jang2009points} is hard to reach, the new target is chosen purely based on the preceding control flow behaviour in contrast with more popular techniques~\cite{ferguson1996chaining}.  
\item Based on the case studies found in two related works~\emph{Confix}~\cite{amin:ase15} and \emph{TAJS}~\cite{dom2011}, we conducted an empirical validation of our framework and performed a significance study of the results. The validation has shown the effectiveness of the framework in covering target branches with a reasonable efficiency. Moreover, we showed that on the selected subjects \emph{JEDI} outperforms \emph{Confix} ....  
\end{enumerate}

The paper is structured as follows: In Section~\ref{sec.example}, we present a motivating example illustrating the challenges of generating test data for JS. The architecture of the \emph{JEDI} testing framework is explained in Section~\ref{sec.framework}. Section~\ref{sec.evaluation} presents empirical validation and indicates the validity threats. Related and future work are discussed in Section~\ref{sec:related.work}. Section~\ref{sec:concl} concludes the paper.

%% The state of JS 2017 survey~\footnote{\url{https://stateofjs.com/}} contains an extensive list of testing frameworks including Mocha, Jasmine, Jest, Enzyme, Tape, Ava, Karma and QUnit. These frameworks support testing at various levels: unit, integration and functional. The survey also indicated a relatively low satisfaction rate of the current state of JS testing tools, scoring only 3.2 out of 5. As result, new testing frameworks such as Cypress and Storybook are constantly appearing to improve the situation. But despite this effort, test data generation continues to be poorly addressed by any of the widely used testing frameworks. The research community has actively been working towards making progress in this direction introducing such tools as Artemis~\cite{artemis2011}, JSConTest~\cite{heidegger2010contract}, Crawljax~\cite{mesbah2012crawling}, JSeft~\cite{mirshokraie2015jseft} and Jalangi~\cite{sen2013jalangi}. A recent study of client-side JS bugs~\cite{ocariza2017study} has shown that the majority (68\%) of web application faults are DOM-related and introduced in the JS code, and not in other components such as HTML or CSS. This fact indicates that yet a better testing approach is needed when it comes to unit testing of JS.

@@

Let us consider the JS function \texttt{isGameFinished} in Figure~\ref{code.isGameFinished} taken from the web game \emph{sudoku}\footnote{\url{http://www.dhtmlgoodies.com/scripts/game_sudoku/game_sudoku.html}}. This function checks the validity of a sudoku solution given by a user. First, we find the DOM element representing the game field (line~\ref{isGameFinished.getSudoku}). Then, we collect all its child DIVs (line~\ref{isGameFinished.getDivs}). Each DIV element corresponds to a sudoku input square (line~\ref{isGameFinished.if1.begin}) and consists of two SPANs. The first span contains a user input, whereas the second one, which is hidden by default, stores the expected value for that sudoku square. Inside of a for-loop in lines~\ref{isGameFinished.inFor.begin}-\ref{isGameFinished.outFor}, we iterate over the whole collection of input squares. Once, we find a square with two unequal SPAN values (line~\ref{isGameFinished.if2.begin}), the loop is terminated and the game is considered unfinished (line~\ref{isGameFinished.unfinished}) and finished otherwise.

Suppose now that we would like to unit test this function in isolation to the rest of the application. We aim to maximize the branch coverage as a common testing criteria~\cite{zhu1997software}. That is, for each individual branch of the function under test (FUT) \texttt{isGameFinished}, we need to construct a test input that covers the branch and leads to the normal termination of the FUT. The JS function takes \emph{explicit} input arguments (in our case none), but it can also accept \emph{implicit} arguments such as the DOM state. Henceforth, we assume that DOM is the only global variable that is allowed to be referred to in the body of the FUT. In general, in order to produce a complete test for our function we have to construct both an appropriate DOM and input arguments.

@@

Figure~\ref{fig.isGameFinished.tests} presents four tests that together provide the full coverage of the FUT. Let us look now at how following the structure of the FUT in Figure~\ref{code.isGameFinished}, we can construct the $T4$ test covering the $(10,11)$ branch. Line~\ref{isGameFinished.getSudoku} expects the input DOM to have an element with the id \textquotesingle\texttt{sudoku}\textquotesingle. In order to enter the for-loop in line~\ref{isGameFinished.inFor.begin}, that element should have at least one child \emph{DIV} tag. The first if-condition in line~\ref{isGameFinished.if1.begin} requires this $div$ to be of the class \textquotesingle\texttt{sudokuSquare}\textquotesingle\ and also have no background color. The second if-condition (line~\ref{isGameFinished.if2.begin}) expects that the $div$ consists of two \emph{SPAN} elements whose \emph{innerHtml} values are not equal. The resulting test $T4$ is shown in Figure~\ref{fig.isGameFinished.tests} and meets all the above conditions.

As we have just seen, even for such a relatively simple example, the construction procedure of a test input is far from trivial because it requires the deep understanding of the program semantics. In practice, web developers often have to deal with a large number of the JS frameworks and libraries freely available on the GitHub whose code is commonly untested and poorly documented. At the same time, unit testing is the most fundamental level of testing and the easiest to implement. It saves the developers from unexpected regressions and often documents the desired behavior of a program.

@@

Given a JS function $f$ and a set of branches $\mathbb{B}_f$, the ultimate goal of the testing framework is to find input arguments for $f$ that cover all branches from $\mathbb{B}_f$. Our JS testing framework, called \emph{JEDI}, is a client-server application which under the hood uses a genetic engine for test generation. The general workflow of the framework is depicted as a sequence diagram in Figure~\ref{fig.framework.architect}. The diagram consists of the four interacting components \User, \Server, \Validator, and \Client. The \Server is the key component of the  framework. It is responsible for the \emph{Initialization Phase} and the \emph{Genetic Phase}. Both phases will be covered in detail in Section~\ref{sub.sec.init.phase} and \ref{sub.sec.genetic.phase} respectively.

Each newly constructed HTML document has to be processed by the \Validator\footnote{\url{https://validator.w3.org/nu/}} which checks its syntactic correctness and reports back the set of identified errors. \Server rejects the document if any errors were revealed and generates a new one until the generated document has no more errors. \Client is a \emph{Node.js}\footnote{\url{https://nodejs.org/en/}} application that is mainly responsible for the execution of the JS code. We use the \emph{jsdom}\footnote{\url{https://github.com/tmpvar/jsdom}} library to model the virtual DOM and simulate the native browser API calls.

@@

Let us take a closer look at the sequence of interactions in Figure~\ref{fig.framework.architect}. As an input for our testing framework, \User has to provide the JS file \emph{fun.js} that contains the function under test (FUT) \emph{fun} annotated with a type signature. Additionally, \User specifies a configuration file \emph{config} that defines various parameters for random and genetic phases, logging etc. Experimentally, we have found an optimal set of input configurations that was used across all experimental objects in Section~\ref{sec.evaluation}. This data is processed by the \Server during the \emph{Initialization Phase}. As an output, we get back the following information about \emph{fun}: the control flow graph (\emph{cfg}), the complete list of its branches (\emph{branches}), the constant info (\emph{cinfo}), the instrumented version \emph{fun} (\emph{ifun}), and the signature (\emph{sig}). The \Server sends a \texttt{POST} request with \emph{ifun} and \emph{sig} to the \Client. Depending on the configured execution mode, the \Server either asks the \User to select a \emph{target} branch or covers all branches one by one in a fixed order. For each branch, our framework enters the \emph{Genetic Loop (Phase)}. Inside of this phase, the framework starts with a randomly generated population (\emph{Random Loop}), and continues to evolve that population (\emph{Fitness Loop}) until either a perfect entity is found or a certain termination criteria is achieved, e.g. the limit on the number of generation attempts. Each candidate in the population has to be evaluated in \emph{Evaluation Loop} by \Client. It matches the candidate $p$ with the type signature \emph{sig} and invokes the instrumented JS function on it: \emph{ifun(p)}. The execution trace along with some supplementary information about the execution are sent back to the \Server, where the \emph{fitness value} is computed. The new population is constructed by combining the results of the \emph{Crossover} and \emph{Mutation Loops}. If a new HTML document is produced somewhere in the genetic phase, it has to be first approved by the \Validator. At the end of the \emph{Fitness Loop}, the candidate with the ``best'' fitness value is reported back to the \User. Once all the branches are covered, the algorithm reaches the end. Below, we elaborate on the details of each individual phase of our test generation framework.

@@

During the \emph{Initialization Phase} we analyze the FUT in order to retrieve the data required for the subsequent genetic testing. The main steps of this phase are described in Algorithm~\ref{alg.init}. Given a FUT \emph{fun.js} as an input, we parse the function and its type signature in line~\ref{alg.init.parse.fun.sig}. In order to maximize the chances of generating the fittest candidate in the genetic phase, in line~\ref{alg.init.collect.info} we perform static analysis of the FUT with the purpose to collect information about constants. Such constants include the numeric and string literals, as well as the arguments to the following DOM API methods: \texttt{getElementsByTagName}, \texttt{getElementById}, \texttt{getElementsByClassName}, and \texttt{getElementsByName}
These methods provide the information about \emph{tags}, \emph{ids}, \emph{classes} and \emph{names} referred from the FUT. The next step is to normalize the FUT before applying any instrumentation (line~\ref{alg.init.transform.ast}), e.g. transform \emph{if-then} branches into \emph{if-then-else}. We instrument the FUT (line~\ref{alg.init.instr}) in order to collect the run-time data such as the execution trace, \emph{the approach level}, \emph{the branch and loop distances} (explained later), and the constants. Upon the \emph{ifun} execution this information is gathered and fed back to the fitness evaluation procedure. Finally, we construct the control-flow graph (line~\ref{alg.init.build.cfg}) and identify all branches of the FUT (line~\ref{alg.init.get.branches}).

@@

\begin{figure}[t!]
\setlength{\grammarparsep}{3pt}
\footnotesize
\begin{grammar}

@@

The grammar of the JS type annotations supported by our test generation framework is shown in Figure~\ref{fig.js.type.annot}. Each type annotation is a signature embedded into a specialized JS type comment (\texttt{/*t...*/}). We distinguish between two kinds of signatures: those which refer to the global DOM and ``ordinary'' JS type signatures. The functions with the former kind expect some \texttt{dom} argument to be passed as the first implicit parameter of the function call. Whereas ordinary JS signature is a non-empty sequence of the JS types. The type can be \emph{Void}, \emph{Primitive} (\texttt{bool}, \texttt{int}, \texttt{float}, \texttt{string}) or \emph{Array} (recursive homogeneous array type).

@@

Figure~\ref{code.isGameFinished.instr} presents the \texttt{isGameFinished} function after instrumentation. The global variable \emph{trace} in line~\ref{isGameFinished.instr.trace} is an array of labels corresponding to the sequence of executed program statements. All statements have to be traced because each one, in theory, can raise an exception. The \texttt{loopMap} map in line~\ref{isGameFinished.instr.loopMap} approximates the potential number of iterations of the respective for-loop. If no exact bound can be computed, we set it to one. The \texttt{branchDistance} map records the conditional distance~\cite{tracey1998automated} on the branch where the deviation from the target branch has happened.

@@

The heart of our test data generation framework is in the \emph{Genetic Phase}. Although, the main workflow of the underlying genetic algorithm (GA) is fairly standard~\cite{poli2008field}, it has been extended with the concept of  \emph{archive convergence} (Algorithm~\ref{alg.gen}). As an input, we take a configuration file \emph{config}, a target branch $b_n$ of the FUT (which we would like to cover), and a tuple of supplementary parameters computed by Algorithm~\ref{alg.init}. The output consists of the ``best'' fitness score $f^*$ achieved by the population candidate $p^*$.

First, we read the \emph{config} file (line~\ref{alg.gen.read.config}) and retrieve the following GA configuration parameters: a population size  ($size_P$), an archive size ($size_A$), the number of generations ($size_G$), crossover ($rate_C$) and mutation ($rate_M$) rates respectively, and a history window size ($size_H$). We start with the initial target $b_n$ (line~\ref{alg.gen.init.target}) and the iteration counter $i$ equal to 0 (line~\ref{alg.gen.iter.count}). We run the genetic search for the target branch (lines~\ref{alg.gen.converged.loop.begin}-\ref{alg.gen.converged.loop.end}) until either the history has not converged, or the generations limit has not been reached. The \emph{archive} variable stores the most fit population candidates observed so far, and the \emph{history} variable records all archives during the generation process. These variables both are empty at the beginning of the evolution (line~\ref{alg.gen.init.archive} and~\ref{alg.gen.init.history} respectively).

The initial population has $size_P$ elements, and it is randomly generated out of the constants from \emph{cinfo} (line~\ref{alg.gen.random.gen}). In the loop~\ref{alg.gen.loop.begin}-\ref{alg.gen.loop.end}, this population is evolved until either a perfect candidate has been found, or the history has converged, or the generation limit $size_G$ has been reached. But first, we need to score the current population (lines~\ref{alg.gen.eval.begin}-\ref{alg.gen.eval.end}). Each candidate $p$ of the population $pop$ represents a concrete substitution for the arguments of the FUT. Thus, it should to be coerced respectively to the type signature $sig$ (line~\ref{alg.gen.make.args}). Afterwards, the instrumented function \emph{ifun} can be called with the coerced arguments $args$ (line~\ref{alg.gen.call.ifun}). This gives us a 4-tuple consisting of an execution trace ($tr$), branch ($d_B$), loop distances ($d_L$), and the set of dynamic constants (\emph{dyncinfo}) collected at the run-time. The next step is actually computing the fitness value for $p$ (line~\ref{alg.gen.eval}). Collected dynamic constants \emph{dyncinfo} are merged with the existing constant pool (line~\ref{alg.gen.eval.end}). Eventually, the scored population is combined with the archive (line~\ref{alg.gen.get.combo}) in $combo$. And it is used for crossover (line~\ref{alg.gen.cross}) and mutation (line~\ref{alg.gen.mut}) steps in order to form the new population (line~\ref{alg.gen.new.pop}). Then, the scored population together with the current archive are sorted by fitness value, and the best $size_A$ elements form the new archive (line~\ref{alg.gen.new.archive}). The element at the top of the archive is the fittest candidate (line~\ref{alg.gen.archive.head}) of the current population. If $f^*$ is a \emph{perfect} candidate, the solution is found (line~\ref{alg.gen.loop.end}). Eventually, the tuple $(f^*, p^*)$ is returned by the algorithm (line~\ref{alg.gen.return}) as solution.

In some cases, when we do not get enough guidance from a fitness function, the genetic algorithm converges to a local minimum~\cite{mcminn2004search}. For example, if one branch is dominated by another, the coverage of the former depends on the coverage of the latter. Unless this relation is encoded in the fitness function somehow, we can not learn about it accidentally. In such a situation, we find ourselves being trapped on a flat space, with a small chance to escape. In order to overcome this problem, we propose a novel --- to the best of our knowledge --- solution based on the idea of \emph{archive convergence}. If during a certain history window, the fitness values of the respective archives have converged to a fixed point, we deem the whole history converged and restart the genetic search with a new search target. At the end of the evolution loop (line~\ref{alg.gen.has.converged}), we verify the convergence of the history within the window of $size_H$. If that condition is true (line~\ref{alg.gen.loop.end}), the genetic algorithm is restarted with a new search target. It consists of the original branch $b_n$ prefixed with a new branch $b$ from which $b_n$ depends on (line~\ref{alg.gen.new.target}). Every time the history converges, we produce a new search target by eventually exhausting all possible executions leading to $b_n$. In practice, we only consider extensions upto length two.

The genetic operations used for the primitive JS types and arrays are summarized in Table~\ref{tbl.gen.oper.js.types}. In the rest of this section, we provide implementation details of random generation in Section~\ref{sub.sub.sec.random.html}, crossover and mutation in Section~\ref{sub.sub.sec.genetic.oper}, and the fitness function in Section~\ref{sub.sub.sec.fitness.fun}. Our main focus here is on the generation and transformation of HTML documents.
